On exponential sums arising from the classical doubling method

### 2025-04-09 main.tex
Typo
Add comment about the factor \chi(2)
Changes with Calvin

\calvin{I think I would like to include a main theorem here even if it's not super precise just to let the reader know what we highlight what we about the sum}
\begin{theorem}[\Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}, \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}]
	For a classical group $G$ and a multiplicative character $\chi$ as above, $\genHermitianJacobiKernel{\hermitianSpace}{\chi}$ are stable functions. We have that
	$$\posDblJacobiSumScalar{\pi}{\chi} =
		\varepsilon_{\algebraicGroup{H}} c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),$$
		and that
	$$\negDblJacobiSumScalar{\pi}{\chi} = \begin{dcases} 0 & G \text{ is an odd special orthogonal group},\\
		\centralCharacter{\pi}\left(-1\right) \posDblJacobiSumScalar{\pi}{\chi} &  \text{otherwise.}
	\end{dcases}$$
Here $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$ is an explicit factor and
$g_T\left(\chi, \theta, \fieldCharacter\right)$ is a product of Gauss sums, see \Cref{subsec:normalization-factor} and \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}.
\end{theorem}
In this paper, we study the doubling method Jacobi sums $\dblJacobiSum{\pi}{\chi}$. We compute the scalars $\dblJacobiSumScalar{\pi}{\chi}$ in terms of the Deligne--Lusztig data corresponding to $\pi$ and show that, up to some well understood factors, they match the Kondo Gauss sum of the conjectural functorial transfer of $\pi$ to a general linear group. This is done by combining a couple of ingredients. The first ingredient is a lemma of Saito--Shinoda \cite{SaitoShinoda2000} that allows us to explicitly compute the analogous character sums for Deligne--Lusztig virtual representations $\RTTheta{T}{\theta}$.  The second ingredient, which is a key ingredient for this computation, is that these Jacobi sums are well behaved, in the sense that they are constant on Lusztig series. To show this, we utilize a multiplicativity property of these Jacobi sums from \cite{GirschZelingher2025} and combine it with a careful analysis of the Deligne--Lusztig data corresponding to $\pi$ after functorial transfer.

@@

Given a torus $T \subset \GL_n\left(\finiteField\right)$ of the form $T \cong \multiplicativegroup{\finiteFieldExtension{\lambda}}$ where $\lambda \vdash n$ is a partition and given characters $\alpha \colon \multiplicativegroup{\finiteFieldExtension{\lambda}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon T \to \multiplicativegroup{\cComplex}$, we define $$\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \left(-1\right)^{\lengthof\left(\lambda\right)} q^{-\frac{n}{2}} \sum_{t \in T} \alpha^{-1}\left(t\right) \chi^{-1}\left(\det t\right) \fieldCharacter\left(\trace t\right).$$
We also define $$\GaussSumSingleTorusCharacter{T}{\alpha}{\fieldCharacter} = \GaussSumTorusCharacter{T}{\alpha}{1}{\fieldCharacter}.$$
Notice that if $\alpha = \alpha_1 \times \dots \times \alpha_{\lengthof\left(\lambda\right)}$, where $\alpha_j \colon \multiplicativegroup{\finiteFieldExtension{\lambda_j}} \to \multiplicativegroup{\cComplex}$ is a character for every $j$, then
\begin{align}
	\label{eq:Gauss_multiplicitive}
	\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \prod_{j=1}^{\lengthof\left(\lambda\right)} \GaussSumCharacter{\alpha_j}{\chi}{\fieldCharacter_{\lambda_j}}.
\end{align}

@@

@@

Let $$w_n = \begin{pmatrix}
	& & & 1\\
	& & 1\\
	& \iddots\\
	1
\end{pmatrix} \in \GL_n\left(\finiteField\right)$$ be the longest Weyl element.
\calvin{we should standardize an order between Unitary, Special Orthogonal Symplectic and try to kinda stick with it throughout.}
\begin{enumerate}
	\item (Special orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. It follows by definition that every element $g \in \IsometryGroup\left(\hermitianSpace\right)$ satisfies $\left(\det g\right)^2 = 1$.  In this case, we obtain the connected components of the identity by adding a restriction on the determinant: We have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det g = 1$. There are two cases here.
	\begin{enumerate}
		\item Split case: in this case, we can write $$\hermitianSpace = \begin{dcases}
		\xIsotropic \oplus \yIsotropic & \dim_{\finiteField} \hermitianSpace = 2n\\
		\xIsotropic \oplus \finiteField v \oplus \yIsotropic & \dim_{\finiteField}\hermitianSpace = 2n + 1
		\end{dcases}$$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In the odd dimensional case, $v \in \hermitianSpace$ is an element orthogonal to $\xIsotropic$ and $\yIsotropic$ that satisfies $\innerproduct{v}{v} \ne 0$.

@@

		In the odd-dimensional case $N = 2n+1$, we are always in this situation, and we simply denote $\SO_{2n+1}\left(\finiteField\right) = \SO_{2n+1}^{+}\left(\finiteField\right)$ and $\algebraicGroup{\SO}_{2n+1} = \algebraicGroup{\SO}_{2n+1}^{+}$.
		\item Non-split case: In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ and there exists a decomposition $$\hermitianSpace = \xIsotropic \oplus \hermitianSpace' \oplus \yIsotropic,$$
		where $\xIsotropic$ and $\yIsotropic$ are totally isotropic spaces in duality and $\hermitianSpace'$ is an antisotropic space (that is, for any $0 \ne v' \in \hermitianSpace'$, we have $\innerproduct{v'}{v'} \ne 0$) of dimension 2. Let $d$ be an element of $\multiplicativegroup{\finiteField}$ that is not a square. Denote $$B_{n,d} = \begin{pmatrix}
			\IdentityMatrix{n-1}\\
			& & 1 &\\
			& -d & &\\
			& & & \IdentityMatrix{n-1}
		\end{pmatrix} \cdot w_{2n}.$$
		In this case, we have that $G$ is isomorphic to $\SO^{-}_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\SO}^{-}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid \transpose{g} B_{n,d} g = B_{n,d},\,\, \det g = 1\right\}.$$
	\end{enumerate}
	\item (Symplectic groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = -1$. In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ for some positive integer $n$ and $G$ is isomorphic to the group $\Sp_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\Sp}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid \transpose{g} \begin{pmatrix}
		& w_n\\
		-w_n
	\end{pmatrix} g = \begin{pmatrix}
		& w_n\\
		-w_n
	\end{pmatrix}\right\}.$$
	\item (Unitary groups): $\quadraticExtension \ne \finiteField$. Denote $\dim_{\quadraticExtension} \hermitianSpace = n$. In this case, the group $G = \GroupExtension{G}$ is isomorphic to the group $\UnitaryGroup_n\left(\finiteField\right) \subset\GL_{n}\left(\quadraticExtension\right)$ defined as the $\finiteField$-points of $$\algebraicGroup{\UnitaryGroup}_n = \left\{ g \in \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{\GL}_n} \mid \involution{\left(\transpose{g}\right)} w_n g = w_n \right\}.$$
\end{enumerate}

@@

The goal of the next two sections is to compute $\dblJacobiSumScalar{\pi}{\chi}$ for $\pi$ that appears in a Deligne--Lusztig virtual representation $\RTThetaVirtual{T}{H}{\theta_H}$ for a suitable maximal torus $T_H$ and $\theta_H \colon T_H \to \multiplicativegroup{\cComplex}$.

@@

Let $\chi \colon \multiplicativegroup{\quadraticExtension} \to \multiplicativegroup{\cComplex} $ be a character. We define a normalization factor as follows $$c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) = \begin{dcases}
	\left(-\GaussSumSingleCharacter{\chi^{-2}}{\fieldCharacter}\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is even,}\\
	\chi\left(2\right) \left(-\GaussSumSingleCharacter{\chi^{-2}}{\fieldCharacter}\right)^{\frac{\dim_{\finiteField} \hermitianSpace - 1}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is odd,}	\\
	\left(-\GaussSumSingleCharacter{\chi^{-1}\restriction_{\multiplicativegroup{\finiteField}}}{\fieldCharacter}\right)^{\dim_{\quadraticExtension} \hermitianSpace} & \quadraticExtension \ne \finiteField.
\end{dcases}$$
Notice that if $\quadraticExtension \ne \finiteField$ then by the Hasse--Davenport relation $\GaussSumSingleCharacter{\chi^{-1}\restriction_{\multiplicativegroup{\finiteField}}}{\fieldCharacter}^{2} = \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}} {\fieldCharacter_{\quadraticExtension}}$. The factor on the left hand side appears in \Cref{thm:computation-unitary-norm-one-case}, while the one on the right hand side appears in \Cref{thm:multiplicativity-in-terms-of-gauss-sums}.

@@

	Hence our theorem follows from the results mentioned above and from the formula $$\RTGTheta{T_H}{H}{\theta_H}\left(1\right) = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} \frac{\grpIndex{H}{T_H}}{\sizeof{H}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{H}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor},$$
	the fact that \begin{equation*}
		\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}},
	\end{equation*}
	and the facts
	\begin{align*}
		\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{1}{2}\dim_{\finiteField} \hermitianSpace \right\rfloor} & \,\,\,\,\,\text{and}\,\,\,\,\, \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}} = \left(-1\right)^r.
	\end{align*}
	In this computation, the factor $\varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}}$ comes from \Cref{thm:split-case-deligne-lusztig-computation} while the factor $\varepsilon_{\algebraicGroup{G}}$ comes from the definition of the normalization factor $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$.
\end{proof}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $H$ such that $\pi$ appears as a subrepresentation of $\RTThetaVirtual{T_H}{H}{\theta_H}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblNegVirtualJacobiSumScalar{\RTGTheta{T_H}{H}{\theta_H}}{\chi} =& \RTTheta{T_H}{H}{\theta_H}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \centralCharacter{\pi}\left(-1\right) \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter}.
	\end{align*}
\end{remark}

@@

	Thus if $\pi$ is a subrepresentation of a parabolically induced representation as above, then $(T_H,\theta_H)$ is geometrically conjugate to $(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H'}, \theta_{\GL_k\left(\quadraticExtension\right)} \times \theta_{H'})$ for $T_{\GL_k\left(\quadraticExtension\right)} \times H'$ a torus of $L = \GL_k\left(\quadraticExtension\right) \times H'$. As proved in \Cref{lem:indep_geo_conj}, $$g_{T \cap G}\left(\chi,\theta_H\restriction_{T \cap G},\psi\right) = g_{\left(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H}'\right) \cap G}\left(\chi,\left(\theta_{\GL_k\left(\quadraticExtension\right)}\times \theta_{H'}\right)\restriction_{\left(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H'}\right) \cap G},\psi\right).$$ The right hand of the last equality is multiplicative since Gauss sums are multiplicative by \eqref{eq:Gauss_multiplicitive}.
\end{proof}
\begin{proof}[Proof of \eqref{item:follows-for-cuspidal}]
By \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters} and \Cref{lem:indep_geo_conj}, we have that if $\RTThetaVirtual{T_H}{H}{\theta_H} = \sum_{\pi} c_{\pi} \pi$, where $\pi$ goes over all (equivalence classes of) irreducible representations of $H$ and $c_{\pi} \in \zIntegers$, then
$$\sum_{\pi} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \RTGTheta{T_H}{H}{\theta_H}\left(1\right) = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \left(\sum_{\pi} c_{\pi}\right).$$
Subtracting the equality for non-cuspidal $\pi$ established by (\ref{item:follows-for-non-cuspidal}), we are left with the equality
$$\sum_{\pi \textnormal{ cuspidal}} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \left(\sum_{\pi \textnormal{ cuspidal}} c_{\pi}\right).$$
But by \Cref{prop:at-most-one-cuspidal-lusztig-series} there is at most one cuspidal $\pi$ appearing in our sum. Thus the proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} is complete.
\end{proof}

>>>>

\calvin{I think I would like to include a main theorem here even if it's not super precise just to let the reader know what we highlight what we about the sum}
\begin{theorem}[\Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig}, \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}]
	For a classical group $G$ and a multiplicative character $\chi$ as above, $\genHermitianJacobiKernel{\hermitianSpace}{\chi}$ are stable functions. We have that
	$$\posDblJacobiSumScalar{\pi}{\chi} =
		\varepsilon_{\algebraicGroup{H}} c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) g_T\left(\chi, \theta, \fieldCharacter\right),$$
		and that
	$$\negDblJacobiSumScalar{\pi}{\chi} = \begin{dcases} 0 & G \text{ is an odd special orthogonal group},\\
		\centralCharacter{\pi}\left(-1\right) \posDblJacobiSumScalar{\pi}{\chi} &  \text{otherwise.}
	\end{dcases}$$
Here $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$ is an explicit factor and
$g_T\left(\chi, \theta, \fieldCharacter\right)$ is a product of Gauss sums, see \Cref{subsec:normalization-factor} and \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}.
\end{theorem}
In this paper, we study the doubling method Jacobi sums $\dblJacobiSum{\pi}{\chi}$. We compute the scalars $\dblJacobiSumScalar{\pi}{\chi}$ in terms of the Deligne--Lusztig data corresponding to $\pi$ and show that, up to some well understood factors, they match the Kondo Gauss sum of the conjectural functorial transfer of $\pi$ to a general linear group. This is done by combining a couple of ingredients. The first ingredient is a lemma of Saito--Shinoda \cite{SaitoShinoda2000} that allows us to explicitly compute the analogous character sums for Deligne--Lusztig virtual representations $\RTThetaVirtual{T}{G}{\theta}$.  The second ingredient, which is a key ingredient for this computation, is that these Jacobi sums are well behaved, in the sense that they are constant on Lusztig series. To show this, we utilize a multiplicativity property of these Jacobi sums from \cite{GirschZelingher2025} and combine it with a careful analysis of the Deligne--Lusztig data corresponding to $\pi$ after functorial transfer.

@@

Given a maximal torus $T \subset \GL_n\left(\finiteField\right)$ of the form $T \cong \multiplicativegroup{\finiteFieldExtension{\lambda}}$ where $\lambda \vdash n$ is a partition and given characters $\alpha \colon \multiplicativegroup{\finiteFieldExtension{\lambda}} \to \multiplicativegroup{\cComplex}$ and $\chi \colon T \to \multiplicativegroup{\cComplex}$, we define $$\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \left(-1\right)^{\lengthof\left(\lambda\right)} q^{-\frac{n}{2}} \sum_{t \in T} \alpha^{-1}\left(t\right) \chi^{-1}\left(\det t\right) \fieldCharacter\left(\trace t\right).$$
We also define $$\GaussSumSingleTorusCharacter{T}{\alpha}{\fieldCharacter} = \GaussSumTorusCharacter{T}{\alpha}{1}{\fieldCharacter}.$$
Notice that if $\alpha = \alpha_1 \times \dots \times \alpha_{\lengthof\left(\lambda\right)}$, where $\alpha_j \colon \multiplicativegroup{\finiteFieldExtension{\lambda_j}} \to \multiplicativegroup{\cComplex}$ is a character for every $j$, then
\begin{align}
	\label{eq:Gauss_multiplicitive}
	\GaussSumTorusCharacter{T}{\alpha}{\chi}{\fieldCharacter} = \prod_{j=1}^{\lengthof\left(\lambda\right)} \GaussSumCharacter{\alpha_j}{\chi}{\fieldCharacter_{\lambda_j}}.
\end{align}

@@

\subsubsection{Algebraic groups}
\elad{Tell the reader that the notation $\algebraicGroup{G}$ refers to an algebraic group, which can be thought of as the $\algebraicClosure{\finiteField}$ points of a variety over $\finiteField$ with an action of $\Frobenius$}

@@

Let $$w_n = \begin{pmatrix}
	& & & 1\\
	& & 1\\
	& \iddots\\
	1
\end{pmatrix} \in \GL_n\left(\finiteField\right)$$ be the longest Weyl element.
\begin{enumerate}
	\item (Unitary groups): $\quadraticExtension \ne \finiteField$. Denote $\dim_{\quadraticExtension} \hermitianSpace = n$. In this case, the group $G = \GroupExtension{G}$ is isomorphic to the group $\UnitaryGroup_n\left(\finiteField\right) \subset\GL_{n}\left(\quadraticExtension\right)$ defined as the $\finiteField$-points of $$\algebraicGroup{\UnitaryGroup}_n = \left\{ g \in \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{\GL}_n} \mid \involution{\left(\transpose{g}\right)} w_n g = w_n \right\}.$$
	\item (Special orthogonal groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = 1$. It follows by definition that every element $g \in \IsometryGroup\left(\hermitianSpace\right)$ satisfies $\left(\det g\right)^2 = 1$.  In this case, we obtain the connected components of the identity by adding a restriction on the determinant: We have that $G$ consists of all elements $g \in \IsometryGroup\left(\hermitianSpace\right)$ with $\det g = 1$. There are two cases here.
	\begin{enumerate}
		\item Split case: in this case, we can write $$\hermitianSpace = \begin{dcases}
		\xIsotropic \oplus \yIsotropic & \dim_{\finiteField} \hermitianSpace = 2n\\
		\xIsotropic \oplus \finiteField v \oplus \yIsotropic & \dim_{\finiteField}\hermitianSpace = 2n + 1
		\end{dcases}$$ where $\xIsotropic$ and $\yIsotropic$ are totally isotropic (that is, for every $x \in \xIsotropic$, $\innerproduct{x}{x} = 0$ and similarly for $\yIsotropic$) and these spaces are in duality with respect to $\innerproduct{\cdot}{\cdot}$, i.e., the map $\yIsotropic \to \Hom_{\finiteField}\left(\xIsotropic, \finiteField\right)$ given by $y \mapsto \left(x \mapsto \innerproduct{x}{y}\right)$ is an isomorphism. In the odd dimensional case, $v \in \hermitianSpace$ is an element orthogonal to $\xIsotropic$ and $\yIsotropic$ that satisfies $\innerproduct{v}{v} \ne 0$.

@@

		In the odd-dimensional case $N = 2n+1$, we are always in this situation, and we simply denote $\SO_{2n+1}\left(\finiteField\right) = \SO_{2n+1}^{+}\left(\finiteField\right)$ and $\algebraicGroup{\SO}_{2n+1} = \algebraicGroup{\SO}_{2n+1}^{+}$.
		\item Non-split case: In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ and there exists a decomposition $$\hermitianSpace = \xIsotropic \oplus \hermitianSpace' \oplus \yIsotropic,$$
		where $\xIsotropic$ and $\yIsotropic$ are totally isotropic spaces in duality and $\hermitianSpace'$ is an antisotropic space (that is, for any $0 \ne v' \in \hermitianSpace'$, we have $\innerproduct{v'}{v'} \ne 0$) of dimension 2. Let $d$ be an element of $\multiplicativegroup{\finiteField}$ that is not a square. Denote $$B_{n,d} = \begin{pmatrix}
			\IdentityMatrix{n-1}\\
			& & 1 &\\
			& -d & &\\
			& & & \IdentityMatrix{n-1}
		\end{pmatrix} \cdot w_{2n}.$$
		In this case, we have that $G$ is isomorphic to $\SO^{-}_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\SO}^{-}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid \transpose{g} B_{n,d} g = B_{n,d},\,\, \det g = 1\right\}.$$
	\end{enumerate}
	\item (Symplectic groups): $\quadraticExtension = \finiteField$ and $\epsilon_{\hermitianSpace} = -1$. In this case, $\dim_{\finiteField} \hermitianSpace = 2n$ for some positive integer $n$ and $G$ is isomorphic to the group $\Sp_{2n}\left(\finiteField\right) \subset  \GL_{2n}\left(\finiteField\right)$, defined as the $\finiteField$-points of $$\algebraicGroup{\Sp}_{2n} = \left\{g \in \algebraicGroup{\GL}_{2n} \mid \transpose{g} \begin{pmatrix}
		& w_n\\
		-w_n
	\end{pmatrix} g = \begin{pmatrix}
		& w_n\\
		-w_n
	\end{pmatrix}\right\}.$$
\end{enumerate}

@@

The goal of the next two sections is to compute $\dblJacobiSumScalar{\pi}{\chi}$ for $\pi$ that appears in a Deligne--Lusztig virtual representation $\RTThetaVirtual{T_H}{H}{\theta_H}$ for a suitable maximal torus $T_H$ and $\theta_H \colon T_H \to \multiplicativegroup{\cComplex}$.

@@

Let $\chi \colon \multiplicativegroup{\quadraticExtension} \to \multiplicativegroup{\cComplex} $ be a character. We define a normalization factor as follows $$c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) = \begin{dcases}
	\left(-\GaussSumSingleCharacter{\chi^{-2}}{\fieldCharacter}\right)^{\frac{\dim_{\finiteField} \hermitianSpace}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is even,}\\
	\chi\left(2\right) \left(-\GaussSumSingleCharacter{\chi^{-2}}{\fieldCharacter}\right)^{\frac{\dim_{\finiteField} \hermitianSpace - 1}{2}} & \quadraticExtension = \finiteField,\,\, \dim_{\finiteField} \hermitianSpace \text{ is odd,}	\\
	\left(-\GaussSumSingleCharacter{\chi^{-1}\restriction_{\multiplicativegroup{\finiteField}}}{\fieldCharacter}\right)^{\dim_{\quadraticExtension} \hermitianSpace} & \quadraticExtension \ne \finiteField.
\end{dcases}$$
Notice that if $\quadraticExtension \ne \finiteField$ then by the Hasse--Davenport relation $\GaussSumSingleCharacter{\chi^{-1}\restriction_{\multiplicativegroup{\finiteField}}}{\fieldCharacter}^{2} = \GaussSumSingleCharacter{\minusInvolutionMinusOne{\chi}} {\fieldCharacter_{\quadraticExtension}}$. The factor on the left hand side appears in \Cref{thm:computation-unitary-norm-one-case}, while the one on the right hand side appears in \Cref{thm:multiplicativity-in-terms-of-gauss-sums}. The factor $\chi\left(2\right)$ in the odd-dimensional special orthogonal case will be explained in the proof of \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}.

@@

	Hence our theorem follows from the results mentioned above and from the formula $$\RTGTheta{T_H}{H}{\theta_H}\left(1\right) = \varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} \frac{\grpIndex{H}{T_H}}{\sizeof{H}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{H}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\finiteField} \hermitianSpace}{2}\right\rfloor},$$
	the fact that \begin{equation*}
		\varepsilon_{\algebraicGroup{H}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{H}}} = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}},
	\end{equation*}
	and the facts
	\begin{align*}
		\varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor \frac{1}{2}\dim_{\finiteField} \hermitianSpace \right\rfloor} & \,\,\,\,\,\text{and}\,\,\,\,\, \varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}} = \left(-1\right)^r.
	\end{align*}
	In this computation, the factor $\varepsilon_{\algebraicGroup{T}_{\algebraicGroup{G}}}$ comes from \Cref{thm:split-case-deligne-lusztig-computation} while the factor $\varepsilon_{\algebraicGroup{G}}$ comes from the definition of the normalization factor $c_{\hermitianSpace}\left(\chi, \fieldCharacter\right)$.
\end{proof}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $H$ such that $\pi$ appears as a subrepresentation of $\RTThetaVirtual{T_H}{H}{\theta_H}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblNegVirtualJacobiSumScalar{\RTGTheta{T_H}{H}{\theta_H}}{\chi} =& \RTGTheta{T_H}{H}{\theta_H}\left(1\right) c_{\hermitianSpace}\left(\chi, \fieldCharacter\right) \centralCharacter{\pi}\left(-1\right) \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter}.
	\end{align*}
\end{remark}

@@

	Thus if $\pi$ is a subrepresentation of a parabolically induced representation as above, then $(T_H,\theta_H)$ is geometrically conjugate to $(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H'}, \theta_{\GL_k\left(\quadraticExtension\right)} \times \theta_{H'})$ for $T_{\GL_k\left(\quadraticExtension\right)} \times T_{H'}$ a torus of $L = \GL_k\left(\quadraticExtension\right) \times H'$. As proved in \Cref{lem:indep_geo_conj}, $$g_{T}\left(\chi,\theta_H\restriction_{T \cap G},\psi\right) = g_{\left(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H}'\right) \cap G}\left(\chi,\left(\theta_{\GL_k\left(\quadraticExtension\right)}\times \theta_{H'}\right)\restriction_{\left(T_{\GL_k\left(\quadraticExtension\right)} \times T_{H'}\right) \cap G},\psi\right).$$ The right hand of the last equality is multiplicative since Gauss sums are multiplicative by \eqref{eq:Gauss_multiplicitive}.
\end{proof}
\begin{proof}[Proof of \eqref{item:follows-for-cuspidal}]
By \Cref{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters} and \Cref{lem:indep_geo_conj}, we have that if $\RTThetaVirtual{T_H}{H}{\theta_H} = \sum_{\pi} c_{\pi} \pi$, where $\pi$ goes over all (equivalence classes of) irreducible representations of $H$ and $c_{\pi} \in \zIntegers$, then
$$\sum_{\pi} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \RTGTheta{T_H}{H}{\theta_H}\left(1\right) = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \left(\sum_{\pi} c_{\pi}\right).$$
Subtracting the equality for non-cuspidal $\pi$ established by (\ref{item:follows-for-non-cuspidal}), we are left with the equality
$$\sum_{\pi \textnormal{ cuspidal}} c_{\pi} \dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = \GaussSumTorusCharacter{T'}{\theta'}{\chi}{\fieldCharacter} \left(\sum_{\pi \textnormal{ cuspidal}} c_{\pi}\right).$$
But by \Cref{prop:at-most-one-cuspidal-lusztig-series} there is at most one cuspidal $\pi$ appearing in our sum. Thus the proof of \Cref{thm:doubling-method-gamma-factor-for-deligne-lusztig} is complete.
\end{proof}
